### 2017-07-18 correspondance/cover.tex
adding cover letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}
 \begin{document}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor D.P. Ames\\
Editor-in-Chief, Environmental Modelling \& Software}

%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\address{Matthew Gidden\\
Laimgrubengasse 17/7\\
Vienna, Austria 1060}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Professor Ames,}

Please find enclosed a manuscript entitled: ``A Methodology and Implementation
of Automated Emissions Harmonization for Use in Integrated Assessment Models''
which I am submitting for exclusive consideration of publication as a
full-length article in Environmental Modelling \& Software.  This manuscript
describes novel work in the subject area of Integrated Assessment Models (IAMs)
and Modelling, specifically related to their use in combination with Earth
Systems Models (ESMs). A new software library, \texttt{aneris}, is presented
which robustly automates the process of emissions trajectory
harmonization. Through harmonization, multiple IAMs can produce results
consistent with a single historical data set.

This capability is critical for the success of the worldwide scientific
community efforts in the Coupled Model Intercomparison Project Phase 6 (CMIP6)
in which results from IAMs are being used as inputs to ESMs in order to model
future scenarios of climate change. By producing a tested, documented, and open
source software implementation, this work

\begin{itemize}
  \item provides a mechanism by which IAMs contributing to the current process
    of CMIP6 can produce consistent results not only within a single team but
    among all teams
  \item allows for other teams in the scientific community to contribute
    consistent results to the CMIP6 process in the future
  \item and increases the transparency and scientfic integrity of that process.
\end{itemize}

Thank you for your consideration of this work. Please address all correspondence
concerning this manuscript to me and feel free to correspond with me by e-mail.

\closing{
Warm Regards,

Matthew Gidden, Ph.D.
}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}