### 2017-07-25 elsfrontmatter.tex
openly available

\begin{frontmatter}

\title{A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models}

%% Group authors per affiliation:
\author[iiasa]{Matthew J. Gidden\corref{corref}}
\ead{gidden@iiasa.ac.at}
\cortext[corref]{Corresponding author}

\author[nies]{Shinichiro Fujimori}
\author[pbl]{Maarten van den Berg}
\author[pik]{David Klein}
\author[pnnl]{Steven J. Smith}
\author[pbl]{Detlef P. van Vuuren}
\author[iiasa]{Keywan Riahi}

\address[iiasa]{International Institute for Applied Systems Analysis,
  Schlossplatz 1, A-2361 Laxenburg, Austria}
\address[nies]{Center for Social and Environmental Systems Research, National Institute for Environmental Studies, 16-2 Onogawa, Tsukuba, Ibaraki 305-8506, Japan}
\address[pbl]{PBL Netherlands Environmental Assessment Agency, Postbus 30314, 2500 GH The Hague, Netherlands}
\address[pik]{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, D-14412 Potsdam, Germany}
\address[pnnl]{Joint Global Change Research Institute, 5825 University Research Court, Suite 3500, College Park, MD 20740}

\begin{abstract}
Emissions harmonization refers to the process used to match greenhouse gas (GHG)
and air pollutant results from Integrated Assessment Models (IAMs) against a
common source of historical emissions. To date, harmonization has been performed
separately by individual modeling teams. For the hand-over of emission data for
the Shared Socio-economic Pathways (SSPs) to climate model groups, a new
automated approach based on commonly agreed upon algorithms was developed. This
work describes the novel methodology for determining such harmonization methods
and an open-source Python software library implementing the methodology. Results
are shown for two example scenarios (with and without climate policy cases)
using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 96\% of the
total emissions trajectories while having a negligible effect on key long-term
climate indicators. This new capability enhances the comparability across
different models, increases transparency and robustness of results, and allows
other teams to easily participate in intercomparison exercises by using the
same, openly available harmonization mechansim.
\end{abstract}

\begin{keyword}
Integrated Assessment Models, Climate Change, Harmonization, Air Pollution 
\end{keyword}

\end{frontmatter}

\linenumbers