### 2018-02-08 paper.tex
adding suggested edits from Keywan

\documentclass[review]{elsarticle}
\usepackage{lineno,hyperref}
\modulolinenumbers[5]
\journal{Environmental Modelling \& Software}
\bibliographystyle{elsarticle-num}
%%%

%%% General usepackages, definitions, etc.
\usepackage[acronym,toc]{glossaries}
\usepackage{graphicx}
\usepackage{booktabs} % nice rules for tables
\usepackage{microtype} % if using PDF
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{moreverb} % for verbatim snippets of code
\usepackage{fancyvrb}
\usepackage{tabularx} % for tables with line breaks
\usepackage{threeparttable} % for tables with notes
\usepackage[capitalize, noabbrev]{cleveref} % for reference multiple figures
\usepackage{calc} % allows for arithmetic on latex variables
\usepackage{algorithm2e} % for algorithms
\usepackage{multirow} % combining rows in tables
\usepackage{footnote}
\usepackage{titlesec} % for using \titleformat
\usepackage{fixltx2e} % for using \textsubscript
\usepackage{bashful}
\usepackage{xspace}
\usepackage{color}
\usepackage{subfig}
\usepackage{minted}
%%%%%

%%%%% Helpers

\newcommand{\code}[1]{\lstinline[basicstyle=\ttfamily\color{black}]|#1|}
\newcommand{\codeb}[1]{\texttt{#1}}
\newcommand{\units}[1] {\:\text{#1}}%
\newcommand{\TODO}[1]{\textbf{TODO: #1}}

%%% gases
\newcommand{\cotwo}{CO\textsubscript{2}~}
\newcommand{\nox}{NO\textsubscript{x}}
\newcommand{\noxx}{NO\textsubscript{x}~}
\newcommand{\nht}{NH\textsubscript{3}}
\newcommand{\nhtt}{NH\textsubscript{3}~}

\makeatletter
\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}
\makeatother

%%% Figure path
\graphicspath{{./figs/}}

%%% Plural gls entries
\newacronym[
  \glslongpluralkey={Integrated Assessment Models},
  \glsshortpluralkey={IAMs}
]{iam}{IAM}{Integrated Assessment Model}
\newacronym[
  \glslongpluralkey={Shared Socioeconomic Pathways},
  \glsshortpluralkey={SSPs}
]{ssp}{SSP}{Shared Socioeconomic Pathway}
\newacronym[
  \glslongpluralkey={Representative Concentration Pathways},
  \glsshortpluralkey={RCPs}
]{rcp}{RCP}{Representative Concentration Pathway}
\newacronym[
  \glslongpluralkey={Greenhouse Gases},
  \glsshortpluralkey={GHGs}
]{ghg}{GHG}{Greenhouse Gas}

%%% Acronyms
\newacronym{sres}{SRES}{Special Report on Emissions Scenarios}
\newacronym{cmip6}{CMIP6}{Coupled Model Intercomparison Project (Phase 6)}
\newacronym{smip}{ScenarioMIP}{Scenario Model Intercomparison Project}
\newacronym{amip}{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
\newacronym{ceds}{CEDS}{Community Emissions Data System}
\newacronym{esm}{ESM}{Earth System Model}
\newacronym{luc}{LULUC}{Land-use and Land-use Change}
\newacronym{weu}{WEU}{Western Europe}
\newacronym{nam}{NAM}{North America}
\newacronym{pao}{PAO}{Pacific OECD}
\newacronym{eeu}{EEU}{Central and Eastern Europe}
\newacronym{fsu}{FSU}{Former Soviet Union}
\newacronym{pas}{PAS}{Other Pacific Asia}
\newacronym{cpa}{CPA}{Centrally-planned Asia and China}
\newacronym{sas}{SAS}{South Asia}
\newacronym{mea}{MEA}{Middle East and North Africa}
\newacronym{afr}{AFR}{Sub-Saharan Africa}
\newacronym{lam}{LAM}{Latin America and the Caribbean}
\newacronym{cli}{CLI}{Command Line Interface}
\newacronym{wmo}{WMO}{World Meteorological Organization}

\makeglossaries

\begin{document}

%%% ELS input
\begin{frontmatter}

\title{A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models}

%% Group authors per affiliation:
\author[iiasa]{Matthew J. Gidden\corref{corref}}
\ead{gidden@iiasa.ac.at}
\cortext[corref]{Corresponding author}

\author[nies]{Shinichiro Fujimori}
\author[pbl]{Maarten van den Berg}
\author[pik]{David Klein}
\author[pnnl]{Steven J. Smith}
\author[pbl]{Detlef P. van Vuuren}
\author[iiasa]{Keywan Riahi}

\address[iiasa]{International Institute for Applied Systems Analysis,
  Schlossplatz 1, A-2361 Laxenburg, Austria}
\address[nies]{Center for Social and Environmental Systems Research, National Institute for Environmental Studies, 16-2 Onogawa, Tsukuba, Ibaraki 305-8506, Japan}
\address[pbl]{PBL Netherlands Environmental Assessment Agency, Postbus 30314, 2500 GH The Hague, Netherlands}
\address[pik]{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, D-14412 Potsdam, Germany}
\address[pnnl]{Joint Global Change Research Institute, 5825 University Research Court, Suite 3500, College Park, MD 20740}

\begin{abstract}
Emissions harmonization refers to the process used to match greenhouse gas (GHG)
and air pollutant results from Integrated Assessment Models (IAMs) against a
common source of historical emissions. To date, harmonization has been performed
separately by individual modeling teams. For the hand-over of emission data for
the Shared Socioeconomic Pathways (SSPs) to climate model groups, a new
automated approach based on commonly agreed upon algorithms was developed. This
work describes the novel methodology for determining such harmonization methods
and an open-source Python software library implementing the methodology. A case
study is presented for two example scenarios (with and without climate policy
cases) using the IAM MESSAGE-GLOBIOM that satisfactorily harmonize over 96\% of
the total emissions trajectories while having a negligible effect on key
long-term climate indicators. This new capability enhances the comparability
across different models, increases transparency and robustness of results, and
allows other teams to easily participate in intercomparison exercises by using
the same, openly available harmonization mechanism.
\end{abstract}

\begin{keyword}
Integrated Assessment Models, Harmonization, Greenhouse Gases (GHGs), Air Pollution 
\end{keyword}

\end{frontmatter}

\linenumbers

%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section*{Software Availability}

\texttt{aneris}, first made available in 2017, is available online at
\url{https://github.com/iiasa/aneris} as a free and open-source Python software
library (approximately 2000 lines of code). The \texttt{aneris} software was
developed by the lead author whose contact information is shown on the title
page of this manuscript. Documentation for the \texttt{aneris} Python package,
including software requirements, is available online at
\url{http://data.ene.iiasa.ac.at/gidden/aneris/}.

\newpage

\section{Introduction}

\glspl{iam} are tools used to understand the complex interactions between
energy, economy, land use, water, and climate systems. \glspl{iam} provide
global projections of systemic change by dividing the world into a number of
representative regions (typically 10 to 30), the definition of which is distinct
for each model \cite{krey_global_2014}. Results from \glspl{iam} are integral in
a number of international studies, which notably include projections of climate
and energy futures. Recently, the \gls{iam} community has developed scenarios
based on the \glspl{ssp} \cite{Riahi2017153} which quantify a variety of
potential global futures. The \glspl{ssp} are designed to be used in research
that include \gls{esm} simulations, climate impact, adaptation and climate
mitigation studies \cite{oneill_new_2013, vuuren_new_2013}.

While \glspl{iam} are implemented in myriad ways\footnote{IAM models are
  numerous and have a long history in the scientific literature. Various IAMs
  have collaborated to produce community IAM documentation (available online:
  \url{http://themasites.pbl.nl/models/advance/index.php/ADVANCE_wiki}) which
  readers can access for a full treatment of model implementation and
  features.}, including simulation and optimization, the core inputs and outputs
are similar across different models. Modeling teams incorporate data on energy
systems, land use, economics, demographics and emissions sources and
concentrations, among other data, in order to provide a consistent starting
point for future projections. The models then provide estimates of future
trajectories of these variables under various socio-economic and technological
assumptions as well as proposed policy constraints, e.g., targets for future
\gls{ghg} emissions.

The emissions trajectories calculated by \glspl{iam} are critical inputs for
ongoing, worldwide scientific community efforts in the \gls{cmip6}
\cite{eyring_overview_2016}, which is utilizing a number of marker \gls{ssp}
scenarios developed by the \gls{iam} community
(\gls{smip}\cite{oneill_scenario_2016},
\gls{amip}\cite{collins_aerchemmip:_2017}, among others). These trajectories are
endogenously calculated by modeling the individual technologies and sectors that
contribute towards the emissions of different air pollutants and \glspl{ghg} as
well as various mitigation technologies. However, the historical emissions
starting points of models can differ by large amounts depending on the region,
sector, and emissions species.

In practice, \glspl{iam} calculate the total source intensity of emitting
technologies, for example the total activity of coal power plants in China, and
incorporate emissions-intensity factors for individual gas species, for example
the quantity of sulfur emissions from coal plants per megawatt-hour of
production. Models are generally \textit{calibrated} to historical data sources
in one or more base years. Results in the historical period may differ between
models as a result of the sometimes large uncertainties in historical data
sets. Models can also differ in their choice of base-year, which may lag behind
available inventory data. In addition, models have varying sectoral, regional,
and fuel aggregations.

The global climate change community has recently developed a new global
historical emissions data set for both anthropogenic emissions (i.e., the
\gls{ceds} \cite{hoesly_historical_2018} and open-burning \gls{luc} emissions
\cite{van_marle_historic_2017}) which, in conjunction with the \gls{ssp}
\gls{iam} trajectories, will be used for climate-related modeling exercises of
\gls{cmip6}.

When participating in intercomparison exercises in which a consistent historical
starting point is required (e.g., in \gls{cmip6}), model teams incorporate a
single, common historical data set through \textit{harmonization}. Harmonization
refers to the process of adjusting model results to match a selected historical
time series such that the resulting future trajectories are consistent with the
original modeled results and provide a smooth transition from the common
historical data. In the emissions context, this means that each individual
combination of model region, model sector, and emissions species must be
harmonized. Depending on the total number of model regions, sectors, and
emissions species, this can require the selection of thousands to
tens-of-thousands of harmonization methods.

Harmonization has been addressed in previous studies as it is a common practice
in the \gls{iam} and climate change communities. For example,
\cite{meinshausen_rcp_2011} describes the use of scaling routines for the 5
regions used in the IPCC \gls{sres} \cite{nakicenovic2000}; however, only total
emissions were harmonized in the exercise, thus there is no sectoral
dimension. Further, \cite{rogelj_discrepancies_2011} describes the impacts of
choosing various harmonization routines on future trajectories. During the
evaluation of the \glspl{rcp}, \gls{iam}
results have been harmonized by sector and the 5 \gls{rcp} global regions
\cite{vuuren_representative_2011}. Importantly, the choice of harmonization
method to date has been determined by individual experts and has generally been
applied to all trajectories for a given class of emissions species.

Climate modeling efforts have continued to progress, demanding increased spatial
and sectoral resolution from \glspl{iam}. Furthermore, a new generation of climate
scenarios which combines aspects of both the \glspl{rcp} and \glspl{ssp} have been developed
in order to incorporate both physical and socio-economic detail. In order to
address the growing dimensionality of model outputs and support ongoing scenario
generation and analysis efforts while still providing a consistent and
scientifically rigorous harmonization procedure, an automated process for
determining harmonization methods is preferred. The use of an automated,
documented, and openly available harmonization mechanism additionally allows for
full procedural reproducibility and for direct participation by additional
modeling teams not involved in the original exercise.

The remainder of this paper describes the methodology and implementation of the
harmonization software \code{aneris} \cite{matthew_gidden_2017_802832}, written
in the Python programming language (detailed documentation is available online
at \url{http://data.ene.iiasa.ac.at/gidden/aneris/}). Section \ref{sec:meths}
provides a detailed description of the underlying mathematical components of
\code{aneris} as well as the procedural workflow. A case study of applying the
automated harmonization mechanism on two example \gls{iam} scenarios, one with
emissions growth and another with emissions mitigation, is presented in Section
\ref{sec:results}. Finally, the general effectiveness and potential future
improvements on the automated methodology is discussed in Section
\ref{sec:future}.

\section{Methodology \& Implementation}\label{sec:meths}

\subsection{The Conceptual Basis for Choosing Harmonization Methods}\label{sec:concept}

The goals of any scenario harmonization exercise are threefold: aligning model
results in the harmonization year to a common historical data source, faithfully
representing the original \gls{iam}’s internal consistency between the driver of
emissions (e.g. energy use) and emissions, and maintaining critical parameters
from the original scenario design. Any harmonization method achieves the first
goal by design. If the difference between the model base year and historical
values are small, considering the second and third goals leads to a method
choice that matches modeled drivers (e.g., a \code{ratio} method discussed in
Section \ref{sec:harm-methods}) and converges prior to the final model year. It
preserves the relationship between \gls{iam} output and emissions inventory in
the base year while also matching the original model output at some point in the
modeled time period. It furthermore maintains the consistency of the model's
usage of energy technology, volume of agricultural activities, and abatement
options with harmonized emissions trajectory.

However, other concerns may lead to a better-informed choice than using a
blanket method for all emissions trajectories. For example, emissions from
\gls{luc} are known to have high year-to-year variation, and therefore
historical data may change drastically depending on the base year considered. In
such a situation, a method that converges at a year past the modeled time period
is a better choice in order to smooth out discrepancies between the historical
data used to develop model and the new data source being used for harmonization.

Separately, if there are large discrepancies between the model results in the
base year and the historic data used for harmonization, convergence methods can
result in harmonized trajectories that do not faithfully represent the
underlying drivers of emissions. Furthermore, if models report negative
emissions, as is possible in scenarios designed to depict the deployment of
climate mitigation policies with large \cotwo reductions and storage, then
end-of-century emissions characteristics should be considered in order to
faithfully match the design parameters of the original scenarios, such as global
mean temperature and other climate metrics.

Accordingly, we have developed a \textit{decision tree} approach to
harmonization method choice, discussed in Section \ref{sec:tree}, in order to
balance each of these concerns and use cases in a robust, systematic,
reproducible, and transparent manner.

\subsection{Harmonization Methods}\label{sec:harm-methods}

\gls{iam} emission results are provided along temporal (normally half decade or
decade), spatial (i.e., model regions), emissions species, and sectoral
dimensions. Each individual temporal trajectory, i.e., unique combinations of
regions ($r$), species ($g$), and sectors ($s$), must be harmonized to the
initial modeling period. Given a model trajectory, $m_{r, g, s}(t)$, historical
values, $h_{r, g, s}(t)$, and model base year, $t_i$, a harmonized trajectory
needs to be calculated. The \textit{harmonization quality} of a trajectory,
i.e., how well a given harmonization algorithm performs, depends on a number of
factors. Of chief import is the faithful representation of the original
unharmonized trajectory as well as the representation of negative trajectories
(i.e., if a trajectory becomes negative, both the timing and total magnitude
should be as close as possible) which are of critical importance for cumulative
\cotwo calculations.

In previous studies \cite{meinshausen_rcp_2011,rogelj_discrepancies_2011}, two
\textit{families} of methods have been used: those that operate on the ratio of
base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and those that operate on the
offset of base year values (i.e., $h(t_i) - m(t_i)$). Both families of functions
depend on a convergence factor, $\beta$, which scales linearly from 1 to 0 over
$[t_i, t_f)$ and is shown in Equation \ref{eqs:factor}. The use of the
  convergence factor implies that the ratio or offset applied in the base year
  ($t_i$) decays to the unharmonized model result (i.e., the convergence factor
  is 0) in the convergence year ($t_f$). In cases where the convergence factor
  is applied over the entire time horizon, the convergence year is taken to be
  $t_f = \infty$.

\begin{equation}\label{eqs:factor}
  \beta(t, t_i, t_f) =
  \begin{cases}
    1 - \frac{t - t_i}{t_f - t_i},& \text{if } t < t_f\\
    0,                        & \text{otherwise}
  \end{cases}
\end{equation}

A number  methods are implemented in \code{aneris} including
ratio-convergence shown in Equation \ref{eqs:ratio}, offset-convergence shown in
Equation \ref{eqs:offset}, and linear interpolation shown in Equation
\ref{eqs:interpolate}. In all equations the region, species, and sector indices
have been dropped for clarity. Each equation is a function of time, model
trajectory, historical trajectory, base year ($t_i$), and a convergence year
($t_f$), at which point the harmonized trajectory converges to the unharmonized
trajectory. \codeb{aneris} provides a number of methods to choose from for each
of the harmonization families. A summary of all available methods is provided in
Table \ref{tab:meths}.

\begin{equation}\label{eqs:ratio}
  m^{rat}(t, m, h, t_i, t_f) = [\beta(t, t_i, t_f) (\frac{h(t_i)}{m(t_i)} - 1) + 1] m(t)
\end{equation}

\begin{equation}\label{eqs:offset}
  m^{off}(t, m, h, t_i, t_f) = \beta(t, t_i, t_f) (h(t_i) - m(t_i)) + m(t)
\end{equation}

\begin{equation}\label{eqs:interpolate}
  m^{int}(t, m, h, t_i, t_f) =
  \begin{cases}
    \beta(t, t_i, t_f) (h(t_i) - m(t_f)) + m(t_f), & \text{if } t < t_f\\
    m(t), & \text{otherwise}
  \end{cases}
\end{equation}

\begin{table}[h!]
\centering
\caption{All harmonization methods provided in \code{aneris}. A Convergence year of $\infty$ is provided for the \code{constant\_ratio} and \code{constant\_offset} methods are listed as $\beta = 1$ for all model years in both cases.}
\label{tab:meths}
\begin{tabular}{|l|l|l|}
\hline
Method Name                             & Harmonization Family & Convergence Year\\
\hline
\code{constant\_ratio}                  & ratio              & $t_f = \infty$\\
\code{reduce\_ratio\_<year>}            & ratio              & $t_f = $\code{<year>}\\
\code{constant\_offset}                 & offset             & $t_f = \infty$\\
\code{reduce\_offset\_<year>}           & offset             & $t_f = $\code{<year>}\\
\code{linear\_interpolate\_<year>}      & interpolation      & $t_f = $\code{<year>}\\
\hline
\end{tabular}
\end{table}

\subsection{Default Method Decision Tree}\label{sec:tree}

A \textit{decision tree} approach has been implemented in \codeb{aneris} which
provides a systematic and documented decision-making process to determine the
preferred harmonization algorithm. In order to provide reasonable
\textit{default} methods, the historical trajectory, unharmonized model
trajectory, and relative difference between history and model values in the
harmonization year are analyzed. The decision tree used in this analysis is a
result of collaborative efforts between \gls{iam} teams and is shown graphically in
Figure \ref{fig:decision_tree}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{decision_tree.pdf}
    \caption[]{
      \label{fig:decision_tree}
      The default method decision tree used in the \code{aneris} software
      library. For all decisions, upper (purple) branches represent a ``yes''
      response and lower (orange) branches represent a ``no'' response. The
      coefficient of variation, $c_v$, is defined in Equation \ref{eqs:cov},
      $dH$ is defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
      decision-making thresholds for $c_v$ and $dH$ are described in the main text.  Where
      present, convergence years of default methods are provided below the
      method name in parentheses.  Convergence years are chosen in order to
      balance the three harmonization goals discussed in Section
      \ref{sec:concept}. Methods labeled in \textit{green} are likely to closely
      match unharmonized results, methods in \textit{yellow} will likely
      somewhat match unharmonized results, and methods in \textit{red} can be
      expected to have a large relative difference between harmonized and
      unharmonized results.}
  \end{center}
\end{figure}

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is rare but possible for a sector to have emissions 
reported in the historical data, but the model to report zero for the harmonization year, 
with non-zero future values. In such cases,
an offset method is required as a ratio method would mask future emissions and
erroneously harmonize the trajectory to zero.

In most cases, however, models do report values in the harmonization
year. Figure \ref{fig:cases} displays a number of example trajectories which
highlight the possible issues resulting from harmonizing model results in
different contexts. These trajectories do not correspond to specific model
results; rather they serve as illustrative examples of the kinds of trajectories
observed in practice. Panels \code{a} and \code{b} highlight examples where
model results peak mid-century, behavior that is seen in a number of scenarios
with general emissions mitigation effects, such as pollution controls applied by
developing nations on transport and industry sectors. Panel \code{a} highlights
a case where model base-year values and history are relative close whereas Panel
\code{b} shows a situation where model values and history are relatively far
apart. Panels \code{c} and \code{d} show similar model trajectories that peak
mid-century but also have negative emissions. Models can report negative
emissions for \cotwo in future scenarios with
climate mitigation enacted via the deployment of carbon capture and
storage technologies. Again, the
relative difference with historical values differ between the panels to explore
harmonization method choices in each situation.

When model and historical values are relatively close, a convergence method is
chosen in order to be as representative as possible to the underlying
unharmonized model results (Figure \ref{fig:cases}, Panel \code{a}). If values
are not close, the constant ratio method is chosen in order to provide
reasonable trajectories that still incorporate modeled effects (Figure
\ref{fig:cases}, Panel \code{b}).

If a model provides a trajectory that transitions from positive to negative
emissions and base year results are similar, then a convergence method is used
in order to guarantee capture of this transition in a representative fashion
(Figure \ref{fig:cases}, Panel \code{c}). If the discrepancy in base year
results is large, it is possible for a negative trajectory to be inappropriately
harmonized to a positive, but decreasing, trajectory. As such, the constant
ratio method is chosen (Figure \ref{fig:cases}, Panel \code{d}).

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{cases.pdf}
    \caption[]{
      \label{fig:cases}
      We present here a number of illustrative examples of the effect of
      different harmonization routines on model trajectories under ``normal''
      circumstances (Panel \code{a}), when there is a large difference between
      historical and model values in the harmonization year (Panels \code{b} and
      \code{d}), and when model trajectories result in negative emissions by the
      end of the modeling time horizon (Panels \code{c} and \code{d}). Identical
      model trajectories are used in each row (Panels \code{a}, \code{b};
      \code{c}, \code{d}). In each column, historical values are increased in
      the base year by an order of magnitude (from 10 to 100). In each Panel, a
      subset of the potential routines provide a better harmonization quality
      than others as described in the text.  }
  \end{center}
\end{figure}

Temporal variability of the historical trajectory is also an important
characteristic when considering the choice of harmonization method.  Emissions
from forest and grassland fires, for example, vary from year to year due to a
combination of meteorological conditions and anthropogenic drivers. Land use
emissions in many \glspl{iam} are modeled using average emission factors and do not
capture conditions in a specific year. A longer convergence horizon is thus
desired in order to incorporate highly variable historical data with modeled
results as is consistency in harmonization method because the effects are
modeled similarly across regions and species. In order to detect emissions with
a high amount of variation, a measure of the coefficient of variation, $c_v$, of
the first derivative of the historical trajectory is calculated using the
standard deviation, $\sigma$, and the mean, $\mu$, as shown in Equation
\ref{eqs:cov}. For a single realization of $c_v$, the first derivative
information of the entire historical time period is utilized.

\begin{equation}\label{eqs:cov}
    c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
\end{equation}

The value of $c_v$ is then tested against a threshold, $\tau_{c_v}$. To
determine this threshold, an analysis of the recent \gls{ceds} and \gls{luc}
historical data has been performed. Figure \ref{fig:cov} shows the distribution
of \gls{luc} $c_v$s and non-\gls{luc} $c_v$s as determined for historical data
aggregated to the model regions of 5 different \glspl{iam} involved in the
\gls{ssp} process: AIM-CGE \cite{fujimori_ssp3:_2017},
IMAGE\cite{van_vuuren_energy_2017}, GCAM4\cite{calvin_ssp4:_2017},
MESSAGE-GLOBIOM\cite{fricko_marker_2017, message_globiom_2016}, and
REMIND-MAGPIE\cite{kriegler_fossil-fueled_2017}, each of which have varying
definitions of native model regions comprising different collections of
countries. Therefore, each data point comprising Figure \ref{fig:cov} represents
a realization of $c_v$ for a single combination of native model region, sector,
and emissions species\footnote{A full listing of all sectors and species is
  presented in the case-study discussion in Section \ref{sec:results}, Table
  \ref{tab:sp}}. A threshold value of $\tau_{c_v} = 20$ has been chosen based on
these observations as it optimally divides the two distributions. Importantly,
tails of the \gls{luc} and non-\gls{luc} overlap, thus there are both false
positives (~7\% of non-\gls{luc} trajectories) and false negatives (~10\% of
\gls{luc} trajectories). However, as any regional definition is model dependent
and thus any regional aggregation is possible an automated detection mechanism
is necessary.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{cov.pdf}
    \caption[]{
      \label{fig:cov}
      The distribution of $c_v$ values for \gls{luc} and non-\gls{luc} historical
      trajectories is shown. \gls{ceds} historical data \cite{hoesly_historical_2018}
      is used for non-\gls{luc} data and \cite{van_marle_historic_2017} is used for
      \gls{luc} data. All historical data has been aggregated from their native
      spatial resolution (i.e., individual countries) to \gls{iam} model regional
      definitions (i.e., collections of countries), and all gas species included
      in the historical data sets are included in the analysis. The solid black
      line indicates the threshold value, $\tau_{c_v}$, used by default in
      \code{aneris}.  }
  \end{center}
\end{figure}

Finally, the default harmonization decisions depend on the relative difference between
the historic and model values in the harmonization time period. In order to
investigate the possible values that these relative differences can take, the
\gls{iam} values used in the \gls{ssp} and (ongoing) \gls{cmip6} inter-comparison
exercises are used. A distribution of these differences for all models in the
study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{dh.pdf}
    \caption[]{
      \label{fig:dh}
      The distribution of relative differences between model and historical
      values in the harmonization year is shown. The solid black line indicates
      the 50\% threshold value, $\tau_{dH}$, used by default in \code{aneris}.
    }
  \end{center}
\end{figure}

\subsection{\code{aneris} Python Implementation and Workflow}\label{sec:workflow}

We herein present \code{aneris}' Python implementation and conceptual
design.. The library is composed of a number of utilities as well as three
primary components: the \texttt{HarmonizationDriver}, \texttt{Harmonizer}, and
data processing routines shown in red, green, and blue, respectively in Figures
\ref{fig:software} and \ref{fig:workflow}.

The \texttt{HarmonizationDriver} is an object designed to interface with
user-provided data and configuration files. Input data (i.e., unharmonized model
results) is assumed to be an Excel file in the standard data format within the
\gls{iam} community, i.e., with \code{Model}, \code{Scenario}, \code{Region},
\code{Variable}, and \code{Unit} columns in addition to columns representing
each modeled time period. It is responsible for down-selecting data into
separate datasets for each model and scenario, invoking the harmonization
process on each dataset, and recompiling the results. The
\texttt{HarmonizationDriver} acts the primary interface for high-level users as
shown by the usage of the \texttt{driver} object in Listing
\ref{lst:tutorial}. Furthermore, a \gls{cli} is provided to allow users to more
easily incorporate the harmonization process in scripted workflows (Listing
\ref{lst:cli}, Figure \ref{fig:software}).

\begin{listing}[H]
\begin{minted}{python}
from aneris.tutorial import load_data
model, hist, driver = load_data()
for scenario in driver.scenarios():
    driver.harmonize(scenario)
harmonized, metadata = driver.harmonized_results()
\end{minted}
\caption{High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://data.ene.iiasa.ac.at/gidden/aneris/tutorial.html}{tutorial}}
\label{lst:tutorial}
\end{listing}

\begin{listing}[H]
\begin{minted}{bash}
$ aneris -h
usage: aneris [-h] [--history HISTORY] [--regions REGIONS] [--rc RC]
              [--output_path OUTPUT_PATH] [--output_prefix OUTPUT_PREFIX]
              input_file

    Harmonize historical trajectories to data in the IAM template format.

    Example usage:

    aneris input.xlsx --history history.csv --regions regions.csv

positional arguments:
  input_file            Input data file.

optional arguments:
  -h, --help            show this help message and exit
  --history HISTORY     Historical emissions in the base year.
  --regions REGIONS     Mapping of country iso-codes to native regions.
  --rc RC               Runcontrol YAML file (see
                        http://data.ene.iiasa.ac.at/gidden/aneris/config.html for
                        examples).
  --output_path OUTPUT_PATH
                        Path to use for output file names.
  --output_prefix OUTPUT_PREFIX
                        Prefix to use for output file names.
\end{minted}
\caption{The \gls{cli} help provided by the \code{aneris} package.}
\label{lst:cli}
\end{listing}

The \texttt{Harmonizer} is a class whose task is to harmonize model value
trajectories given historical data and possible user method \textit{overrides},
i.e., non-default methods (described further in Section \ref{sec:overrides}). It
is used by the \texttt{HarmonizationDriver}; however it is also available to the
user as a first-class object. The \texttt{Harmonizer} requires that input data
conform to the \code{aneris} calculation data format, which explicitly separates
the emissions species from the sector contributing the emissions (these are
combined in the single \code{Variable} column in the standard \gls{iam}
format). Because the \texttt{Harmonizer} is designed to operate on a single
instance of a model and scenario, the canonical data format includes
\texttt{region}, \texttt{sector}, \texttt{gas}, and \texttt{units} columns
without extraneous meta-data columns for the model and scenario. Once configured
with appropriate input data (model and history) as well as potential method
overrides, the \texttt{Harmonizer}'s \texttt{harmonize()} method can be invoked
which returns a \texttt{pandas.DataFrame} of harmonized data. The object can
additionally be queried directly as to its \texttt{default\_methods()},
\texttt{methods()} (i.e., methods used with overrides), and \texttt{metadata()}
(i.e., methods used with all branching information along each path in the
decision tree).

There are also a variety of tools and utilities provided to users and also used
by the \texttt{HarmonizationDriver} in order to process both input and output
data. These include an \texttt{EmissionsAggregator} class and related routines
used to generate sectoral emissions totals, generate regional totals, and
combine historical emissions to native model regions (where historical data is
defined at a higher spatial resolution than a model; see, e.g., Figure
\ref{fig:regions}). A \texttt{FormatTranslator} class is also provided which
defines an interface for translating \texttt{pandas.DataFrame}s between the \gls{iam}
format expected for input and output data and the calculation format used by
\code{aneris}' \texttt{Harmonizer}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{implementation_schematic.pdf}
    \caption[]{
      \label{fig:software}
      The various objects and their relation to one another in the \code{aneris}
      code base as well as a short description of their scope of concern.  
    }
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[height=0.8\textheight]{aneris-workflow.pdf}
    \caption[]{
      \label{fig:workflow}
The full harmonization process as executed by \code{aneris} for a single
instance of a model and scenario. Operations that can be configured with
user-based input configurations are shown in purple. Operations governed by the
\texttt{HarmonizationDriver} are shown in red. Data processing operations are
shown in blue. The core harmonization process, governed by the
\texttt{Harmonizer} is shown in green. }
  \end{center}
\end{figure}

%% \begin{figure}
%%   \subfloat[Python Implementation\label{fig:software}]{
%%     \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
%%     }
%%   \subfloat[Workflow\label{fig:workflow}]{
%%     \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
%%     }
%%     \caption{ Panel \ref{fig:software} shows the various objects in the
%%       \code{aneris} code base as well as a short description of their scope.
%%       Panel \ref{fig:workflow} displays the full harmonization process
%%       as executed by \code{aneris} for a single instance of a model and
%%       scenario. Operations that can be configured with user-based input
%%       configurations are shown in purple. Operations governed by the
%%       \texttt{HarmonizationDriver} are shown in red. Data processing operations
%%       are shown in blue. The core harmonization process, governed by the
%%       \texttt{Harmonizer} is shown in green.  }
%%     \label{fig:software-workflow}
%% \end{figure}

The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins
by cleaning input data. Cleaning operations include adding model
trajectories with 0 values where a sector/emission combination exists in the historical data set but are not provided by the
model input and detecting any issues that would cause the harmonization process
to fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
spatial aggregation to common analysis regions is performed. For example, the
5-region aggregation developed in the \glspl{rcp}
\cite{vuuren_representative_2011} process is commonly used in the \gls{iam}
community and is shown in Figure \ref{fig:regions}. Finally, any exogenous
trajectories the user provides are added. Exogenous trajectories are normally
provided for unmodeled gases with well-accepted scenario trajectories, e.g.,
chlorofluorocarbons provided by the \gls{wmo} \cite{wmo2014}. Upon completion,
the harmonized trajectories and meta data regarding the harmonization process
are returned. A description of all returned meta data is provided in Table
\ref{tab:metadata}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{MESSAGE_11-5regions_map.pdf}
    \caption[]{
      \label{fig:regions}
      The 5 regions used in the \glspl{rcp} with their MESSAGE-GLOBIOM 11-region
      constituents: Asia (\gls{cpa}, \gls{sas}, \gls{pas}) [yellows], \gls{lam}
      [magenta], Africa and the Middle East (\gls{mea} and \gls{afr}) [greens],
      the OECD (\gls{nam}, \gls{weu}, and \gls{pao}) [blues], and the Reforming
      Economies (\gls{eeu} and \gls{fsu}) [reds].  }
  \end{center}
\end{figure}

\begin{table}[]
\centering
\caption{Meta data provided by the \code{aneris} harmonization routine. This meta data is provided for every combination of region, sector, and emissions species.}
\label{tab:metadata}
\begin{tabular}{|p{2cm}|p{8cm}|}
\hline
\textbf{Column}       & \textbf{Description}                                      \\
\hline
\hline
method       & The harmonization method used.                                               \\
\hline
default      & The default harmonization method as determined by the default decision tree. \\
\hline
override     & The method provided as an override (if any).                                 \\
\hline
offset       & The offset value between history and model in the harmonization year.        \\
\hline
ratio        & The ratio value between history and model in the harmonization year.         \\
\hline
cov          & The coefficient of variation value of the historical trajectory.                           \\
\hline
unharmonized & The unharmonized value in the harmonization year.                            \\
\hline
history      & The historical value in the harmonization year.                             \\
\hline
harmonized   & The resulting harmonized value in the harmonization year.\\
\hline
\end{tabular}
\end{table}

\subsection{User-Defined Override Methods}\label{sec:overrides}

Users are able to control the harmonization process via a number of options
(with examples provided
\href{http://data.ene.iiasa.ac.at/gidden/aneris/config.html}{online}). The primary mechanism
by which users control the process is by providing \textit{override} methods for
any combination of region and variable (i.e., sector and gas species).  In
practice, it may be possible that not all default methods chosen will provide
robust harmonized trajectories, especially if there is a significant difference
between historical and model values in the harmonization year, if there is
significant upward or downward movement in the model trajectory, or if there are
known discrepancies in sectoral definition between the \gls{iam} and historical data
source. In such cases, users can override default methods with their method of
choice and both the default method and override is reported in the resulting
metadata.

In order to help identify cases where overrides may be needed, harmonization
\textit{diagnostics} are provided which report the relative difference between
harmonized and unharmonized trajectories at their mid and end-points when these
values exceed specified thresholds. The diagnostic reporting thresholds are
configurable by the user, but defaults of $400$\% and $200$\%, respectively, are
provided based on experiences of the authors' use of \code{aneris} to date.

\section{Case Study: Harmonizing Results from a Global IAM}\label{sec:results}

In order to show a representative cross section of the performance of the
\code{aneris} harmonization procedure, we focus on the harmonization of results
of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017,
  message_globiom_2016}. Harmonization results for two scenarios from the
\gls{ssp} scenario library\footnote{We refer the reader to the broad literature
  discussing the SSPs, e.g., \cite{Riahi2017153, oneill_new_2013,
    vuuren_new_2013,oneill_scenario_2016} for a more in depth discussion of the
  scenario architecture and design. Model results for various SSP scenarios are
  available online at \url{https://tntcat.iiasa.ac.at/SspDb}} are presented
here. We use the SSP2-reference\cite{fricko_marker_2017,Rao2017346}, or ``middle
of the road'', scenario (referred to as SSP2-Ref) as an example because
MESSAGE-GLOBIOM is the marker scenario\footnote{The ``marker'' scenario concept
  is used to designate the archetype scenario used as a reference within each
  SSP scenario family. See \cite{Riahi2017153} and \cite{vuuren_new_2013} for a
  more lengthy description.} for this \gls{ssp}. This SSP2 scenario lies between
two \glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing
  in this context is the energy imbalance at the top of the atmosphere caused by
  anthropogenic influences relative to a pre-industrial reference point. Higher
  radiative forcing leads to larger global changes, such as surface
  temperature. We refer the reader to \cite{IPCCWG1PhysicalStocker2013} for a
  more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
additionally present the results for the SSP2-based mitigation scenario leading
to a radiative forcing of 4.5 Wm$^{-2}$ (referred to as SSP2-4.5). The SSP2-45
scenario is chosen because mitigation technologies and policies are enacted
causing a general reduction in pollutants and \glspl{ghg}, including (eventual)
negative \cotwo emissions in some regions and sectors due to carbon capture and
sequestration and afforestation. A scenario in which negative emissions play a
role in mitigation strategies is particularly important because of the
sensitivity of key indicators, such as end-of-century radiative forcing (which
is used to estimate mean global temperature response), to the timing and
magnitude of net-zero and total negative \cotwo emissions. Therefore, these two
scenarios represent two contrasting cases in the use of a harmonization approach
and thus provide a case study as to its general applicability. Figure
\ref{fig:kyoto} shows the different trends of Kyoto Gases, a measure of
aggregate \gls{ghg} emissions, in each scenarios.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_kyoto.pdf}
    \caption[]{
      \label{fig:kyoto}
      Unharmonized global Kyoto gas emissions for SSP2-Ref, a scenario with
      generally increasing global emissions trends, and SSP2-45, a scenario with
      generally decreasing global emissions trends.  }
  \end{center}
\end{figure}

MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
mapped directly to the 5-region definition used in the RCPs (Figure
\ref{fig:regions}); harmonization is performed directly on the native regional
spatial resolution. Historical data is taken from previously described \gls{luc}
and anthropogenic sources, which comprise 10 separate pollutant and \gls{ghg}
species and 12 sectors shown in Table \ref{tab:sp}. A total of 970 distinct
trajectories\footnote{Table \ref{tab:sp} compiles 24 global trajectories and 86
  regional trajectories. Therefore, with 11 model regions, 970 total
  trajectories are harmonized.} were harmonized for each scenario, and therefore
1940 trajectories were harmonized in total for these two illustrative scenarios.
\noxx generated from the Energy sector provides an example of an emissions
species and sector in which all regions were satisfactorily harmonized with the
default methods. Figure \ref{fig:nox} shows the results of harmonization in
Asia, and Table \ref{tab:nox} describes the parameters that underlie the choice
of method for each harmonized trajectory. Importantly, the default methods
provide regional trajectories for both the reference and mitigation scenarios
that match the general unharmonized model behavior and also result in global
aggregated emissions that are similar to the unharmonized trajectory.

\begin{table}[]\begin{minipage}{\textwidth}
  \centering
  \caption{Harmonized Species and Sectors}
  \label{tab:sp}
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Emissions Species}        & \textbf{Sectors}                    \\
    \hline
    \hline
    Black Carbon (BC)                 & Agricultural Waste Burning\footnoteref{x_co2}          \\
    Hexafluoroethane (C\textsubscript{2}F\textsubscript{6})
    \footnote{\label{x_glb_all}
      Global total trajectories are harmonized due to lack of detailed 
      historical data.}
                                      & Agriculture\footnoteref{x_co2}                         \\
    Tetrafluoromethane (CF\textsubscript{4})
    \footnoteref{x_glb_all}               & Aircraft
    \footnote{\label{x_glb}
      Global sectoral trajectories are harmonized due to lack of detailed 
      historical data.}                            \\
    Methane (CH\textsubscript{4})                     & Energy Sector                       \\
    Carbon Dioxide (CO\textsubscript{2})
    \footnote{\label{x_co2}
      A global trajectory for land-use CO\textsubscript{2} is used; non-land-use sectors 
      are harmonized for each model region.}
                                      & Forest Burning\footnoteref{x_co2}                      \\
    Carbon Monoxide (CO)              & Grassland Burning\footnoteref{x_co2}                   \\
    Hydrofluorocarbons (HFCs)
    \footnoteref{x_glb_all}               & Industrial Sector                   \\
    Nitrous Oxide (N\textsubscript{2}O)
    \footnoteref{x_glb_all}
                                      & International Shipping\footnoteref{x_glb}              \\
    Ammonia (\nht)                     & Residential Commercial Other        \\
    Nitrogen Oxides (\nox)             & Solvents Production and Application \\
    Organic Carbon (OC)               & Transportation Sector               \\
    Sulfur Hexafluoride (SF\textsubscript{6})
    \footnoteref{x_glb_all}               & Waste                               \\
    Sulfur Oxides (SO\textsubscript{x})               &                                     \\
    Volatile Organic Compounds (VOCs) &                                     \\
    \hline
  \end{tabular}
\end{minipage}\end{table}

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{example_NOx_Energy_Sector.pdf}
    \caption[]{
      \label{fig:nox}
      \noxx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
      trajectories for SSP2 and SSP2-45 with historical trajectories (grey
      lines) are presented. The SSP2 reference scenario is shown in Panels
      \textbf{a} and \textbf{c}; the SSP-45 scenario is denoted with ``x''
      markers in Panels \textbf{b} and \textbf{d}. The upper panels show the
      results for endogenously modeled and harmonized regions in Asia while the
      lower panels display the aggregate region results.  
    }
  \end{center}
\end{figure}

\begin{table}[]
\centering
\caption{Key Parameters for Deciding Harmonization Methods for \noxx Emissions in the Energy Sector in Asia}
\label{tab:nox}
\begin{tabular}{|p{1.2cm}|p{.5cm}|p{.5cm}|p{4.5cm}|p{3cm}|}
\hline
\textbf{Region} & \textbf{dH} & $\mathbf{c_v}$ & \textbf{Decision Tree Traversal (Branch and Direction)} & \textbf{Default Method Chosen} \\ \hline
    \hline
\gls{cpa}             & 0.35        & 2.26          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
\gls{pas}             & 0.14        & 1.24          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
\gls{sas}             & 0.56        & 0.58          & 1 (no), 2 (no), 3 (no), 4 (no)                          & constant\_ratio        \\ \hline
\end{tabular}
\end{table}

The harmonization of emissions pathways is performed in order to accurately
represent new or updated data sets of historical emissions inventories while also
maintaining consistency with the original, unharmonized pathway. As such, when
the default methods as provided by the harmonization procedure distort or
otherwise sufficiently misrepresent the underlying unharmonized results, an
override method is required to be provided for the trajectory of the region,
sector, and species in question. Of the 970 trajectories, approximately 10\%
were reported as a diagnostic (see Section \ref{sec:workflow}) of which 3.5\%
required the use of harmonization overrides after an initial investigation;
thus, 96.5\% of all trajectories were satisfactorily harmonized using the
default methods. The trajectories that required overrides clustered into two
classifications: regional trajectories whose \textit{magnitude} was overly
distorted and regional trajectories whose \textit{shape} was overly distorted.

Figure \ref{fig:co} presents a case in which the magnitude of a trajectory is
distorted. A large discrepancy ($\sim$300\% relative difference) is observed in
the harmonization year for carbon monoxide (CO) emissions in the industrial
sector specifically for the \gls{sas} and \gls{afr} regions. Furthermore,
emissions in both regions see relatively large expansion or contraction,
depending on the scenario; therefore, both regions are candidates for choosing
harmonization method overrides. The default method chosen in both cases
(\code{constant_ratio}) maintains model trends for the region; however, overall
model results are distorted as can be seen in both the regional and global
panels in Figure \ref{fig:co}. By applying \code{constant_offset} overrides, the
regional trends and magnitudes are maintained. The use of harmonization
overrides also ameliorates issues seen in the harmonized global trajectory of
industrial CO (Figure \ref{fig:co}, Panel \textbf{c}). Not only does the global
trajectory with overrides more closely match the original unharmonized model
behavior and magnitude of emissions, but the relative importance of the
underlying regional trajectories is also maintained.

\begin{figure}
  \begin{center}
    \includegraphics[height=0.8\textheight]{sep_regions_results_CO_Industrial_Sector.pdf}
    \caption[]{
      \label{fig:co}
      CO Industrial Sector harmonized (without markers) and unharmonized (with
      markers) emissions are presented for SSP2 (purple lines) and SSP2-45 (red
      lines) scenarios. The left column shows the default harmonization and the right column shows harmonization with overrides. Panel \textbf{a} shows  trajectories for the \gls{sas} region
      while panels \textbf{b} and \textbf{c} show  trajectories for the
      \gls{afr} region and (total) global emissions, respectively. Notably, the
      \gls{sas} and \gls{afr} regional trajectories are distorted
      when the default methods are used due to the harmonization-year difference between history and
      model results in both scenarios. The distortion is large enough to affect
      global results, as shown in Panel \textbf{c}.  The use of overrides (right column) results in better consistency with the unharmonized model scenario at the regional and global levels. }
  \end{center}
\end{figure}

In certain circumstances, the application of the default harmonization methods
can affect not only the magnitude but also the shape of regional
trajectories. Figure \ref{fig:nh3} shows an example case of emissions
trajectories for ammonia (\nht) from the agriculture sector. Again, the
\gls{sas} region shows a large discrepancy in the harmonization year ($>$150\%
in this case). The resulting trajectory harmonized with the default method
(\code{constant_ratio}) provides a large increase after 2080 in the SSP2
reference scenario. Notably, the SSP2-45 scenario is not affected to the same
degree. While this distortion changes the magnitude of the \gls{sas} trajectory,
it largely affects the post-2080 shape of the global trajectory (Figure
\ref{fig:nh3}, Panel \textbf{b}) as well as the relative regional contributions
to the global aggregate trajectory. For example, in the original model result,
\gls{sas} \nht~ agricultural emissions contribute $\sim$30\% of total global
emissions, whereas in the harmonized case with default methods, \gls{sas}
comprises $\sim$50\% of global emissions by 2100. By using a
\code{constant_offset} method as an override, this distortion is addressed and
more accurately reflects unharmonized results in the \gls{sas} region, the
relative importance between regions, and global results for agricultural ammonia
emissions, each of which contributes to a better harmonization quality for the
harmonized \gls{sas} trajectory. The harmonized emissions using overrides result
in a contribution of $\sim$32\% of total global emissions by the \gls{sas}
region which aligns closely with the unharmonized model results.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{sep_regions_results_NH3_Agriculture.pdf}
    \caption[]{
      \label{fig:nh3}
      \nhtt agricultural harmonized and unharmonized emissions are presented for
      SSP2 and SSP2-45 scenarios. Panel \textbf{a} shows harmonized and
      overridden-harmonized (respectively) trajectories for the \gls{sas}
      region, and panel \textbf{b} shows the global total trajectory. In this
      case, the \gls{sas} trajectory again shows not only a magnitude distortion, but
      also a shape distortion at the tail of the trajectory. Additionally,
      global trajectories are greatly affected by the harmonization method
      choice (there is $\sim$20\% relative difference between trajectories in
      the reference scenario in 2100). Override methods have been applied to
      correct the distortion.  }
  \end{center}
\end{figure}

We investigate the aggregate effect of harmonization with all necessary override
methods on total anthropogenic radiative forcing projections with the simple
carbon-cycle and climate model, MAGICC\footnote{MAGICC is a reduced complexity
  climate model which incorporates future trajectories of forcing agents (i.e.,
  emissions) to estimate future radiative forcing and mean global temperature
  response. We refer the reader to the MAGICC wiki (available online:
  \url{http://wiki.magicc.org}) for a more in-depth description.}
\cite{meinshausen2011emulating, meinshausen2011rcp}, for each harmonized and
unharmonized scenario respectively as shown in Figure \ref{fig:forcing}. We find
that the change due to harmonization is small, ranging between 1 and 2.5\% over
the modeled time period. Relative near-term differences persist in the
mitigation case (SSP-4.5) because differences in near-term emissions define to a
larger degree the longer-term forcing outcome due to the cumulative nature of
long-lived climate forcers like \cotwo.  The resulting difference in forcing in
2100 is 0.04 $\frac{\text{W}}{\text{m}^2}$ for SSP2-4.5 and 0.01
$\frac{\text{W}}{\text{m}^2}$ for SSP2-Ref, both of which are well within
acceptable tolerances (e.g., 0.75 $\frac{\text{W}}{\text{m}^2}$ defined for
\gls{smip} \cite{oneill_scenario_2016}). Thus harmonization is considered to
have a negligible effect on key long-term climate indicators.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_forcing.pdf}
    \caption[]{
      \label{fig:forcing}
      The results of the simple climate model, MAGICC6, forced with the SSP2-Ref
      (blue) and SSP2-4.5 (green) harmonized and unharmonized scenarios is
      presented. The radiative forcing trajectories of harmonized and
      unharmonized scenarios are shown in solid lines and dashed lines,
      respectively.  }
  \end{center}
\end{figure}

\section{Discussion \& Future Work}\label{sec:future}

This work presented a novel methodology and Python implementation of automated
emissions harmonization for \glspl{iam}, \code{aneris}. An in-depth explanation
of the processes and methods for determining the use of harmonization methods
was provided in Section \ref{sec:meths}. \code{aneris} was able to
satisfactorily harmonize over 96\% of the 1940 individual trajectories that
were analyzed in Section \ref{sec:results}. Of the remaining trajectories,
harmonization method overrides were applied, and  discussion was provided
detailing why overrides were deemed necessary.

The automated approach drastically reduces the need for expert opinion in
determining harmonization methods for each individual combination of model
region, sector, and emissions species while still providing a justifiable
explanation for each automated choice of harmonization method based on both the
historical and future emissions trajectories. Furthermore, the automated
approach should continue to scale well as models become more detailed in both the
regional and sectoral dimensions. Finally, expert opinion is still allowed to
trump the automated method as determined by the algorithm via method overrides;
however, these cases are clearly documented via the meta data provided as an
output of \code{aneris} and thus can be individually explained. This provides
not only transparency and reproducibility, but also scientific integrity in the
choice of harmonization methods.

The use of an open-source, automated harmonization process also provides
benefits to the wider climate science and \gls{iam} communities. By providing a
standard mechanism for harmonization, the \gls{iam} community can directly provide
input into the harmonization algorithms and rules for their default
selection. Additionally, modeling teams are easily capable of executing
identical harmonization procedures in order to participate in ongoing and
further iterations of intercomparison exercises and analysis. Future scenario
analyses can also utilize this common harmonization approach such that they are
consistent with prior efforts.

There are a variety of avenues for future improvement of both the \code{aneris}
software and underlying methodology. As with any software project, additional
users will provide use cases for more robust handling of input/output issues
and corner cases. Further configuration parameters can also be added in the
future in order to provide overrides for all gas species in a given sector or
region. Perhaps the most fruitful investigation will involve further refinement
of the default decision tree introduced in Section \ref{sec:meths}. A key aspect
missing from the decision tree is input from models regarding whether missing
sources or activity levels are the likely cause of a harmonization year
discrepancy (suggesting the use of an offset method) or instead a significant
difference in emissions factors (suggesting the use of a ratio method)
\cite{rogelj_discrepancies_2011}.

This work provides a new direction and framework which the \gls{iam} and climate
science communities can build upon in order to reduce the necessity of
consistent expert opinion and increase transparency and reproducibility of
harmonization exercises. Furthermore, it provides an open-source, tested, and
documented software library which can be used and improved upon by these
communities. Both of these are clear steps in a positive direction for future
climate and integrated assessment modeling exercises.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}

The authors would like to acknowledge a number of colleagues who helped
contribute both discussion and feedback regarding this work including Drs. Elmar
Kriegler, Gunnar Luderer, and Joeri Rogelj. This project has received funding
from the European Union’s Horizon 2020 research and innovation programme under
grant agreement No 641816. The authors further wish to thank the Global
Environment Facility for their generous financial support.

\newpage
%Print the glossary
\printglossaries

\newpage
\section*{\refname}
\bibliography{refs}
\end{document}